---
title: "Stat 5000 Final Project"
author: "Siri D"
date: "11/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(repr)
library(randomForest)
library(rpart)
library(caret)
library(performance)
library(e1071)
library(caTools)
```

## Load the data into R studio
```{r}
heart = read_csv("C:/Users/SIRIDEVA/Documents/UCB/STAT 5000/heart.csv")
glimpse(heart)
```

The data contains the following columns-

1. age - Age in Years
2. sex - (1 = male; 0 = female)
3. cp - Chest Pain Type
4. trestbps - Resting Blood Pressure (in mm Hg on admission to the hospital)
5. chol - Serum Cholesterol (in mg/dl)
6. fbs - Fasting Blood Sugar > 120 mg/dl (1 = true; 0 = false)
7. restecg - Resting Electro Cardiograph (ECG) Results
8. thalach - Maximum Heart Rate Achieved
9. exang - Exercise Induced Angina (1 = yes; 0 = no)
10. oldpeak - ST Depression Induced by Exercise Relative to Rest
11. slope - The Slope of the Peak Exercise ST Segment
12. ca - Number of Major Vessels (0-4) Colored by Flourosopy
13. thal - (1 = normal; 2 = fixed defect; 3 = reversible defect)
14. target - Have a Heart Disease or Not (1 = yes, 0 = no)

We can change the column names into a more understandable form: 

```{r}
colnames(heart)<-c("age","sex","chest_type","resting_bp","chol","fasting_bp",
                   "rest_ecg","max_heart_rate","excersize_induced_angina",
                   "old_peak","slope","num_major_vessels","thal","heart_disease")
glimpse(heart)

heart_data <- heart
```
We store heart data into another tibble heart_data as we wil manipulate the heart table in the future.

### How are the variables correlated to each other?
```{r fig.width=13}
library(corrr)
heart.cor = corrr::correlate(heart_data)
heart.cor %>% rplot(colours = c("red", "green")) 
```


### Is there a particular age where heart diseases occur the most ?

```{r fig.width=8,fig.height=5}

heart$Agecat1<-cut(heart$age, c(10,20,30,40,50,60,70,80,90))
heart%>% ggplot()+
  geom_bar(aes(x=factor(Agecat1)),stat="count", width=0.7, fill = "blue")+ 
  xlab("Age categories")+
  theme_minimal()+ facet_wrap(facets = heart$heart_disease) 
```
Most people who have heart diseases are in their forties and fifties.


###Does cholestrol contribute to heart diseases, if so which age group has the highest cholestrol levels?

```{r fig.width=8,fig.height=6}
heart %>% ggplot( aes(x=Agecat1, y=chol, fill=heart_disease)) + xlab("Age categories")+ 
  ylab("Cholestrol")+geom_bar(stat="identity", position=position_dodge())+ 
  facet_wrap(facets = heart$heart_disease)

```
People who are in their sixties who have cholestrol levels more than 400 are more likely to have a heart disease.

### Age is a normal distribution, let us see if this holds true for people who have heart diseases too.
```{r}
heart_age <- (heart %>% filter(heart_disease == 1))$age
mean(heart_age)
sd(heart_age)
y_dist <- dnorm(heart_age,mean =mean(heart_age),sd =sd(heart_age))
plot(heart_age,y_dist)
```
The ages of people who have heart diseases also follows a normal distribution with mean of 52 years.

### Does resting bp indicate heart diseases, if so which range of heart rates is the most common?
```{r fig.width=11}
heart$Resting_bp_cat<-cut(heart$resting_bp, c(94,105,116,127,138,149,160,171,182,192))
heart%>% filter(na.rm = TRUE)%>% ggplot()+
  geom_histogram(aes(x=factor(Resting_bp_cat),fill = heart_disease),stat="count", width=0.7)+ 
  xlab("Resting bp Categories")+theme_minimal()+ facet_wrap(facets = heart$heart_disease)+
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90))
```
Most heart disease patients have resting bp rates in the range of 127 - 138.

### Does heart rate indicate heart diseases, if so which range of heart rates is the most common?
```{r fig.width=9,fig.height=11}
heart$thalach_cat<-cut(heart$max_heart_rate, c(70,84,97,111,123,137,150,163,176,190))
heart%>% filter(na.rm = TRUE)%>% ggplot()+
  geom_histogram(aes(x=factor(thalach_cat),fill = heart_disease),stat="count", width=0.7)+ 
  xlab("Maximum heart rate Categories")+
  theme_minimal()+ facet_wrap(facets = heart$heart_disease,nrow = 2)

```
Most heart disease patients have maximum heart rates in the range of 150 - 163.

### Which gender is more likely to have a heart disease?
```{r fig.width=8,fig.height=5}
sex <- as_tibble(c("female","male"),c(96,207))
sex<- sex %>% add_column(n = c(207,96))
ggplot(sex, aes(x="", y = n, fill = value)) +
  geom_bar(stat="identity", width=1) +coord_fixed() +
  coord_polar("y",start = 0)
```

### Can most heart diseases be cured?
```{r fig.width=8,fig.height=5}
heart_rate_defect <- as_tibble(c("unknown","normal","fixed defect","reversible defect"))
heart_rate_defect<- heart_rate_defect %>% add_column(n = c(2,18,166,117))
ggplot(heart_rate_defect, aes(x="", y = n, fill = value)) +
  geom_bar(stat="identity", width=1) +coord_fixed() +
  coord_polar("y",start = 0)
```
From the pie chart we can see that most heart diseases are reversible.

### How do fasting bp change with resting bp ?
```{r fig.width=9}
heart %>% group_by(age)%>% ggplot(aes(x = age))+
  geom_line(aes(y = fasting_bp),color = "darkred")+ geom_line(aes(y = resting_bp ),color = "blue")

```
The fasting bp is pretty much the same but the resting bp fluctuates.

### What is the most common type of chest pain?
```{r fig.width=8,fig.height=5}
heart_rate_defect <- as_tibble(c(0,1,2,3))
heart_rate_defect<- heart_rate_defect %>% add_column(n = c(39,41,69,16))
ggplot(heart_rate_defect, aes(x="", y = n, fill = value)) +
  geom_bar(stat="identity", width=1) +coord_fixed() +
  coord_polar("y",start = 0)
```
Type 2 chest pain is the most common.

```{r}
x <- c("age","sex","chest_type","resting_bp","chol","fasting_bp","rest_ecg","max_heart_rate",
       "excersize_induced_angina", "old_peak","slope","num_major_vessels","thal")
y <- heart["heart_disease"]
```

### Let us split the data into training and test to implement models. 
```{r}
train_data <- heart_data[1:250,]
test_data <- heart_data[251:303,]
```

### Linear regression
```{r}
model <- glm(heart_disease ~.,family=binomial(link='logit'),data=train_data)
summary(model)
```
Chest pain, resting blood pressure, number of major vessels blocked and maximum heart rate have high influence over a person having a heart disease.

```{r}
performance_accuracy(model)
```

### Normalise the data for future models.
```{r}
train_scaled <- scale(train_data[,-1])
test_scaled <- scale(test_data[,-1])

```

### SVM
```{r}
classifier <- svm(formula = heart_disease ~ .,
                 data = train_data, scale = TRUE,
                 kernel = 'linear')
summary(classifier)
```


```{r}
pred = predict(classifier, test_data,type = response)
pred <- floor(0.5 + pred)
cm <- table(pred, test_data$heart_disease)
accuracy <- c(28/53)
accuracy
```

### Decision tree
```{r}
fit <- rpart(heart_disease~., data = train_data, method = 'class')
pred <- predict(fit, test_data, type = 'class')
table_mat <- table(test_data$heart_disease, pred)
table_mat
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)

summary(fit)
accuracy_Test
```

### Random forest
```{r}
rf <- randomForest(heart_disease~., data=train_data, proximity=TRUE)
p1 <- predict(rf, train_data)
p1 <- floor(0.5 + p1)
rf_table <- table(p1, train_data$heart_disease)
accuracy_Test <- sum(diag(rf_table)) / sum(rf_table)
accuracy_Test
```


### Which model has the highest accuracy 
```{r}
models <- c('linear regression','svm','decision tree', 'random forest')
accuracy_values <- c(90.32,52,62, 99.2)
ggplot()+ geom_bar(aes(x = models, y = accuracy_values, fill = accuracy_values),stat = 'identity')
```

Random Forests have an accuracy of 99.2%. 

### References 
* https://www.datacamp.com/community/tutorials/support-vector-machines-r
* https://www.r-bloggers.com/2021/04/random-forest-in-r/
* www.datatechnotes.com
* https://www.guru99.com/r-decision-trees.html
* https://www.statology.org/r-glm-predict/